API design for machine learning software: experiences from the scikit-learn project

### 2013-08-05 paper.tex
Rewording

The library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a contiguous numeric array datatype
and fast array computing primitives,
while SciPy \citep{varoquaux2013scipy} extends it
further with common numerical operations, either by implementing
these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
implementations. Building upon this stack, a series of libraries called
\textit{scikits} were created, to complement SciPy with
domain-specific toolkits. Currently, the two most popular and feature-complete
ones are by far \sklearn and
\textit{scikit-image}\footnote{\url{http://scikit-image.org}},
which does image processing.

@@

\citet{pedregosa2011} briefly presented \sklearn and
benchmarked it against several competitors.
In this paper, we instead present an
in-depth analysis of design choices made when building the library,
detailing how we organized and operationalized
common machine learning concepts.
We first present in section~\ref{sec:core-api} the central application
programming interface (API) and then describe, in section~\ref{sec:advanced-api},
advanced API mechanisms built on the core interface.
Section~\ref{sec:implementation} briefly describes the implementation.
Section~\ref{sec:comparison} then
compares \sklearn to other major projects in terms of API\@.
Section~\ref{sec:future_work} outlines some of the objectives and goals for
a \sklearn 1.0 release.
We conclude by summarizing the major points of this paper in
section~\ref{sec:conclusions}.

@@

Actual learning is performed by the \texttt{fit} method. This method is called
with training data (e.g. supplied as two arrays \texttt{X\_train} and
\texttt{y\_train} in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data and set these as attributes on the estimator object. As a convention, the
parameters learned by an estimator are exposed as public attributes with names
suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
learned coefficients of a linear model),
again to facilitate model inspection.
In the partial application view,
\texttt{fit} is a function from data to a model of that data.
As such, it always returns the estimator object it was called on,
which now serves as a model of its input and can be used to perform predictions or transformations of input data.
The choice to let a single object serve dual purpose as estimator and model
simplifies the \sklearn class hierarchy
compared to an approach where estimators and models are separated.

@@

Building on the core interface introduced in the previous section, we now
present advanced API mechanisms for building meta-estimators,
composing complex estimators and selecting models. We also discuss design
choices allowing for an easy use and extension of \sklearn.

@@

A distinguishing feature of the \sklearn API is its ability to
compose new estimators from several base estimators. Composition mechanisms can
be used to combine typical machine learning workflows into a single object which
is itself an estimator, and can be employed wherever usual estimators can be used.
In particular,
they integrate within model selection routines, allowing one to optimize at once
every step of a complex workflow. Composition of estimators can be done in two
ways: either sequentially through \texttt{Pipeline} objects or in a parallel
fashion through \texttt{FeatureUnion} objects.

>>>>

The library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a contiguous numeric array datatype
and fast array computing primitives,
while SciPy \citep{varoquaux2013scipy} extends it
further with common numerical operations, either by implementing
these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
implementations. Building upon this stack, a series of libraries called
\textit{scikits} were created, to complement SciPy with
domain-specific toolkits. Currently, the two most popular and feature-complete
ones are by far \sklearn and
\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
which does image processing.

@@

\citet{pedregosa2011} briefly presented \sklearn and
benchmarked it against several competitors.
In this paper, we instead present an
in-depth analysis of design choices made when building the library,
detailing how we organized and operationalized
common machine learning concepts.
We first present in section~\ref{sec:core-api} the central application
programming interface (API) and then describe, in section~\ref{sec:advanced-api},
advanced API mechanisms built on the core interface.
Section~\ref{sec:implementation} briefly describes the implementation.
Section~\ref{sec:comparison} then
compares \sklearn to other major projects in terms of API\@.
Section~\ref{sec:future_work} outlines some of the objectives for
a \sklearn 1.0 release.
We conclude by summarizing the major points of this paper in
section~\ref{sec:conclusions}.

@@

Actual learning is performed by the \texttt{fit} method. This method is called
with training data (e.g., supplied as two arrays \texttt{X\_train} and
\texttt{y\_train} in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data and set these as attributes on the estimator object. As a convention, the
parameters learned by an estimator are exposed as public attributes with names
suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
learned coefficients of a linear model),
again to facilitate model inspection.
In the partial application view,
\texttt{fit} is a function from data to a model of that data.
As such, it always returns the estimator object it was called on,
which now serves as a model of its input and can be used to perform predictions or transformations of input data.
The choice to let a single object serve dual purpose as estimator and model
simplifies the \sklearn class hierarchy
compared to an approach where estimators and models are separated.

@@

Building on the core interface introduced in the previous section, we now
present advanced API mechanisms for building meta-estimators,
composing complex estimators and selecting models. We also discuss design
choices allowing for easy usage and extension of \sklearn.

@@

A distinguishing feature of the \sklearn API is its ability to
compose new estimators from several base estimators. Composition mechanisms can
be used to combine typical machine learning workflows into a single object which
is itself an estimator, and can be employed wherever usual estimators can be used.
In particular,
they integrate within model selection routines, allowing one to optimize
every step of a complex workflow at once. Composition of estimators can be done in two
ways: either sequentially through \texttt{Pipeline} objects or in a parallel
fashion through \texttt{FeatureUnion} objects.
